Online resource 4. Supplementary graphs and detailed descriptions of XRD and SEM-EDS results

### 2025-04-23 analysis/supplementary-materials/ESM_4.qmd
ESM_4 finished. Rendered version not included due to size.

---
title: "Title Goes Here"
author:
  - Jane Doe:
      correspondence: "yes"
      email: janedoe@fosg.org
      orcid: 0000-0003-1689-0557
      institute:
        - fosg
        - fop
  - John Q. Doe:
      institute: fosg
      orcid: 0000-0003-1689-0558
  - Peder Ås:
      institute: fosg
      orcid: 0000-0003-1689-0559
  - Juan Pérez:
      orcid: 0000-0003-1689-0551
      institute:
        - name: Acme Corporation
  - Max Mustermann:
      orcid: 0000-0003-1689-0552
institute:
  - fosg:
      name: Formatting Open Science Group
      address: 23 Science Street, Eureka, Mississippi, USA
  - fop: Federation of Planets
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/SOM"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
---

# Online resource 4

## XRD results

In this subsection we present the individual spectra graphs for all XRD samples, organised by geological formation, excluding those presented in the main paper and the results analysed and interpreted at the Geociencias Barcelona-CSIC. The original data necessary to produce these visualisations can be found and freely accessed in our online compendium, in the associated GitHub project under the path "../data/xrd_data".

@@

library(readr)
library(dplyr)
library(ggplot2)
library(janitor)
library(tidyr)
library(ggpmisc)

@@

# Read grouped supporting data
xrd_perc <- read_csv('analysis/data/raw_data/xrd_data/xrd_goals.csv') # File with component percentage of all samples for annotations
xrd_marks <- read_csv('data/xrd_data/xrd_peaks.csv') # File with peak data and d spacing for plot labels

@@

# Read individual xrd files
sp6_xrd <- read_csv('data/xrd_data/SP6_xrd.csv')
sp7_xrd <- read_csv('data/xrd_data/SP7_xrd.csv')
sp10_xrd <- read_csv('data/xrd_data/SP10_xrd.csv')
sp14_xrd <- read_csv('data/xrd_data/SP14_xrd.csv')
sp18_xrd <- read_csv('analysis/data/raw_data/xrd_data/SP18_xrd.csv')
sp21_xrd <- read_csv('data/xrd_data/SP21_xrd.csv')
sp28_xrd <- read_csv('data/xrd_data/SP28_xrd.csv')
sp30_xrd <- read_csv('data/xrd_data/SP30_xrd.csv')
sp33_xrd <- read_csv('data/xrd_data/SP33_xrd.csv')
sp34_xrd <- read_csv('data/xrd_data/SP34_xrd.csv')
sp40_xrd <- read_csv('data/xrd_data/SP40_xrd.csv')
sp50_xrd <- read_csv('data/xrd_data/SP50_xrd.csv')
sp52_xrd <- read_csv('data/xrd_data/SP52_xrd.csv')
sp57_xrd <- read_csv('data/xrd_data/SP57_xrd.csv')
sp59_xrd <- read_csv('data/xrd_data/SP59_xrd.csv')
sp61_xrd <- read_csv('data/xrd_data/SP61_xrd.csv')
sp62_xrd <- read_csv('analysis/data/raw_data/xrd_data/SP62_xrd.csv')
sp63_xrd <- read_csv('data/xrd_data/SP63_xrd.csv')
sp66_xrd <- read_csv('data/xrd_data/SP66_xrd.csv')

@@

```{r}
#| label: fig-xrd-sp6
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Cabo S. Vicente (SP6_CSV). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

@@

sp6_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 5000) |> 
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp6_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp6_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp6_perc)))

@@

```{r}
#| label: fig-xrd-sp7
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Cabo S. Vicente (SP7_CSV). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

@@

sp7_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  rename(Quartz = `Quartz (SiO2)`) |> 
  rename(Dolomite = `Dolomite (CaMg(CO3)2)`) |> 
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 65000)) +
  geom_point(data = sp7_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp7_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 60000, label = list(cbind(sp7_perc)))

@@

```{r}
#| label: fig-xrd-sp10
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Ponta dos Altos (SP10_PdA). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

@@

sp10_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 55000)) +
  geom_point(data = sp10_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp10_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp10_perc)))

@@

```{r}
#| label: fig-xrd-sp14
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Praia do Belixe (SP14_PBX). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

@@

sp14_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Dolomite_nb, -Halite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite, Halite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Halite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp14_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp14_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1", "darkgoldenrod1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp14_perc)))

@@

```{r}
#| label: fig-xrd-sp18
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Praia do Belixe (SP18_PBX). The rhombus shape signifies the main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

@@

sp18_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Dolomite_nb, -Halite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite, Halite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Halite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp18_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp18_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1", "darkgoldenrod1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp18_perc)))

@@

```{r}
#| label: fig-xrd-sp21
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Belixe Sul, north of Praia do Belixe (SP21_BLS). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

@@

sp21_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 55000)) +
  geom_point(data = sp21_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp21_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp21_perc)))

@@

```{r}
#| label: fig-xrd-sp28
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Aspa, north of Cabo S. Vicente (SP28_ASP). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

@@

sp28_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 60000)) +
  geom_point(data = sp28_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp28_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp28_perc)))

@@

```{r}
#| label: fig-xrd-sp30
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Foz dos Fornos (SP30_FZF). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

@@

sp30_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 60000)) +
  geom_point(data = sp30_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp30_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp30_perc)))

@@

```{r}
#| label: fig-xrd-sp33
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Foz dos Fornos (SP33_FZF). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

@@

```{r}
#| label: fig-xrd-sp34
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Ponta dos Altos (SP34_PdA_a). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

@@

```{r}
#| label: fig-xrd-sp40
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Ferrel (SP40_b). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

@@

```{r}
#| label: fig-xrd-sp50
#| fig-cap: Plot of minerals identified through XRD in the Middle Jurassic (Malhão formation) chert sample from the outcrop Casal da Colina (SP50_CdC). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

@@

### Geological samples

>>>>

---
title: "Online resource 4. Supplementary graphs and detailed descriptions of XRD and SEM-EDS results"
subtitle: "Within chert: a multi-technique mineral and geochemical approach to the study of chert of southwestern Iberia"
author: Joana Belmiro
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
    toc: true
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
---

# XRD results

In this section we present the individual spectra graphs for all XRD samples, organised by geological formation, excluding those presented in the main paper and the results analysed and interpreted at the Geociencias Barcelona-CSIC. The original data and code necessary to produce these visualisations can be found and freely accessed in our online compendium (https://doi.org/10.17605/OSF.IO/W78V6), in the associated GitHub project under the path "../data/raw_data/xrd_data" and "../supplementary-materials/ESM_4", respectively.

@@

library(readr)
library(dplyr)
library(ggplot2)
library(janitor)
library(tidyr)
library(ggpmisc)
library(here)

@@

# Read grouped supporting data
xrd_perc <- read_csv(here::here('analysis/data/raw_data/xrd_data/xrd_goals.csv')) # File with component percentage of all samples for annotations
xrd_marks <- read_csv(here::here('analysis/data/raw_data/xrd_data/xrd_peaks.csv')) # File with peak data and d spacing for plot labels

@@

# Read individual xrd files
sp6_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP6_xrd.csv'))
sp7_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP7_xrd.csv'))
sp10_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP10_xrd.csv'))
sp14_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP14_xrd.csv'))
sp18_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP18_xrd.csv'))
sp21_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP21_xrd.csv'))
sp28_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP28_xrd.csv'))
sp30_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP30_xrd.csv'))
sp33_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP33_xrd.csv'))
sp34_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP34_xrd.csv'))
sp40_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP40_xrd.csv'))
sp50_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP50_xrd.csv'))
sp52_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP52_xrd.csv'))
sp57_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP57_xrd.csv'))
sp59_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP59_xrd.csv'))
sp61_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP61_xrd.csv'))
sp62_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP62_xrd.csv'))
sp63_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP63_xrd.csv'))
sp66_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP66_xrd.csv'))

@@

```{r}
#| label: fig-xrd-sp6
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Cabo S. Vicente (SP6_CSV). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.
#| fig-width: 6
#| fig-asp: 0.6
#| fig-align: center

@@

sp6_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 5000) |> 
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity (count)") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 60000)) +
  geom_point(data = sp6_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp6_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp6_perc)))

@@

```{r}
#| label: fig-xrd-sp7
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Cabo S. Vicente (SP7_CSV). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.
#| fig-width: 6
#| fig-asp: 0.6
#| fig-align: center

@@

sp7_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  rename(Quartz = `Quartz (SiO2)`) |> 
  rename(Dolomite = `Dolomite (CaMg(CO3)2)`) |> 
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity (count)") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 65000)) +
  geom_point(data = sp7_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp7_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 60000, label = list(cbind(sp7_perc)))

@@

```{r}
#| label: fig-xrd-sp10
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Ponta dos Altos (SP10_PdA). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.
#| fig-width: 6
#| fig-asp: 0.6
#| fig-align: center

@@

sp10_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity (count)") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 55000)) +
  geom_point(data = sp10_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp10_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp10_perc)))

@@

```{r}
#| label: fig-xrd-sp14
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Praia do Belixe (SP14_PBX). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.
#| fig-width: 6
#| fig-asp: 0.6
#| fig-align: center

@@

sp14_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Dolomite_nb, -Halite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite, Halite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Halite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity (count)") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp14_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp14_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1", "darkgoldenrod1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp14_perc)))

@@

```{r}
#| label: fig-xrd-sp18
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Praia do Belixe (SP18_PBX). The rhombus shape signifies the main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.
#| fig-width: 6
#| fig-asp: 0.6
#| fig-align: center

@@

sp18_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Dolomite_nb, -Halite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite, Halite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Halite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity (count)") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 58000)) +
  geom_point(data = sp18_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp18_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1", "darkgoldenrod1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp18_perc)))

@@

```{r}
#| label: fig-xrd-sp21
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Belixe Sul, north of Praia do Belixe (SP21_BLS). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.
#| fig-width: 6
#| fig-asp: 0.6
#| fig-align: center

@@

sp21_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity (count)") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 55000)) +
  geom_point(data = sp21_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp21_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp21_perc)))

@@

```{r}
#| label: fig-xrd-sp28
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Aspa, north of Cabo S. Vicente (SP28_ASP). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.
#| fig-width: 6
#| fig-asp: 0.6
#| fig-align: center

@@

sp28_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity (count)") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 60000)) +
  geom_point(data = sp28_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp28_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp28_perc)))

@@

```{r}
#| label: fig-xrd-sp30
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Foz dos Fornos (SP30_FZF). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.
#| fig-width: 6
#| fig-asp: 0.6
#| fig-align: center

@@

sp30_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity (count)") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 60000)) +
  geom_point(data = sp30_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp30_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp30_perc)))

@@

```{r}
#| label: fig-xrd-sp33
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Foz dos Fornos (SP33_FZF). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.
#| fig-width: 6
#| fig-asp: 0.6
#| fig-align: center

@@

```{r}
#| label: fig-xrd-sp34
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Ponta dos Altos (SP34_PdA_a). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.
#| fig-width: 6
#| fig-asp: 0.6
#| fig-align: center

@@

```{r}
#| label: fig-xrd-sp40
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Ferrel (SP40_b). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.
#| fig-width: 6
#| fig-asp: 0.6
#| fig-align: center

@@

```{r}
#| label: fig-xrd-sp50
#| fig-cap: Plot of minerals identified through XRD in the Middle Jurassic (Malhão formation) chert sample from the outcrop Casal da Colina (SP50_CdC). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.
#| fig-width: 6
#| fig-asp: 0.6
#| fig-align: center

@@

## Geological samples
